---
title: "R Spatial Assignment"
author: "Zachary Kuloszewski"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 80), tidy = TRUE)
```

## set paths
```{r}
rm(list=ls())

user_path <- function() {
  # Return a hardcoded path that depends on the current user, or the current 
  # working directory for an unrecognized user. If the path isn't readable,
  # stop.
  
user <- Sys.info()["user"]
if (user == "zachkuloszewski"){
    path = "/Users/zachkuloszewski/Library/CloudStorage/Dropbox/My Mac (Zachs-MBP.lan)/Documents/GitHub/phd_psets/year2/development/pde/"
  } 
  else if (user == ""){
    path = ""
  } 
  else {
    warning("No path found for current user (", user, ")")
    path = getwd()
  }
  stopifnot(file.exists(path))
  return(path)
}
```


# Problem 1: Mapping Dell (2010)

### Part 1.i, load necessary libraries
```{r, message=FALSE}
library(tidyverse)
library(sf)
library(raster)
library(rgdal)
library(exactextractr)
```

### Part 1.ii-iv, importing raw files
```{r, message=FALSE, warning=FALSE}
mita_shape  <- st_read(paste0(user_path(), 
                              "data/Dell_raw_data_files/MitaBoundary.shp"))
peru_border <- st_read(paste0(user_path(), 
                              "data/Dell_raw_data_files/peru_nw.shp"))
capitals    <- st_read(paste0(user_path(), 
                              "data/Dell_raw_data_files/locations.shp"))
```

### Part 1.v, recast district capitals
```{r}
capitals <- st_transform(capitals, crs(mita_shape))
```

### Part 1.vi, plot a map! 
```{r}
# define map theme
map_theme <- list(theme_bw(),
                  theme(line = element_blank(), 
                        panel.grid.minor=element_blank(), # blank background
                        panel.grid.major=element_blank(),
                        axis.text.x = element_blank(),
                        axis.text.y = element_blank(),
                        axis.ticks = element_blank(), # turn off ticks 
                        axis.title.x = element_blank(), # turn off titles 
                        axis.title.y = element_blank()))

# plot map
peru_map <- ggplot() + 
            geom_sf(data=peru_border) +
            geom_sf(data=capitals, color="blue", size=0.3) +
            geom_sf(data=mita_shape, color = "red") +
            map_theme + 
            ggtitle("Peru, Mitas, and Capital Locations")

peru_map

```

### Part 1.vii, distance to border
```{r}
# calculate distance to both borders
capitals$border_dist <- st_distance(capitals$geometry, mita_shape$geometry)

# take minimum of two distances
capitals <- capitals %>% 
            rowwise() %>%
            mutate(border_dist = min(border_dist[,1], border_dist[,2]))
```

### Part 1.viii, import raster of nightlights
```{r}
nightlights <- raster(paste0(user_path(), "data/per_viirs_100m_2012.tif"))
```

### Part 1.ix, nightlights within 5km of capitals
```{r, message=FALSE}
# calculate buffer objects
capital_buffer <- st_buffer(capitals$geometry, dist=5000)

# mean within each buffer
capitals$mean_nightlights <- exact_extract(nightlights,
                                           capital_buffer, 
                                           'mean')
```

### Part 1.x, mean nightlights in mita capitals
```{r}
# within the mita
mean(capitals$mean_nightlights[capitals$mita == 1 & !is.na(capitals$mita)])

# outside the mita
mean(capitals$mean_nightlights[capitals$mita == 0 | is.na(capitals$mita)], 
     na.rm=T)
```

### Part 1.xi, map everything now
```{r}
# define bounds in latitude and longitude to zoom in on study area
disp_win_wgs84 <- st_sfc(st_point(c(-74.5,-16.5)), st_point(c(-70.5,-13)),
                         crs = 4326)

# switch to right CRS
disp_win_wgs84 <- st_transform(disp_win_wgs84, crs = crs(capitals))
disp_win_coord <- st_coordinates(disp_win_wgs84)

capitals$ln_nightlights <- log(.1+capitals$mean_nightlights)

# plot map
nightlight_map <- ggplot() +
                  geom_sf(data=peru_border) +
                  geom_sf(data=capitals, aes(color=ln_nightlights), size=0.7) +
                  geom_sf(data=mita_shape, color = "black") +
                  map_theme + 
                  ggtitle("Capitals and Nightlights") +
                  coord_sf(xlim=disp_win_coord[,'X'], 
                           ylim=disp_win_coord[,'Y'],
                           datum=crs(capitals), 
                           expand=FALSE) +
                  scale_color_gradient(low="red", high="green")


nightlight_map
```

# Problem 2: African Ethnic Groups

### Part 2.i-ii, load data
```{r, message=FALSE}
# murdock map
tribal_borders <- st_read(paste0(user_path(), "data/Africa/borders_tribes.shp"))

# explorer routes
explorer_rtes  <- st_read(paste0(user_path(), "data/Africa/Explorer_Routes_Final.shp"))
```

### Part 2.iii, selecting pre-colonial voyages
```{r}
explorer_rtes_pre_col <- filter(explorer_rtes, Year_End <= 1885 & !is.na(Year_End))
```

### Part 2.iv, identify ethnic groups with pre-colonial contact
```{r}
# first identify the intersections
tribal_borders$explorers_contacted <- st_intersects(tribal_borders$geometry,
                                                    explorer_rtes_pre_col$geometry)

# then create binary based on length of intersection list
tribal_borders <- tribal_borders %>% 
                  rowwise() %>% 
                  mutate(contacted = as.numeric(length(explorers_contacted) > 0))
```

### Part 2.v, map contacted groups
```{r}
# factorise contact variable for legend
tribal_borders$contacted <- as.factor(tribal_borders$contacted)

contact_map <- ggplot() +
               geom_sf(data=tribal_borders, aes(fill=contacted)) +
               geom_sf(data=explorer_rtes_pre_col) +
               map_theme + 
               ggtitle("Pre-Colonial Exploration and Ethnic Group Contact")
              
contact_map
```

